\bf Eye Fitting Straight Lines in the Modern Era

### 2022-02-04 Eye-Fitting-Straight-Lines-in-the-Modern-Era.Rmd
just a few more edits to go.

Much of the research on the perception of visual features in charts has been conducted in psychophysics and tests for accuracy and quantitative comparisons when understanding a plot. 
Cleveland and McGill conducted a series of cognitive tasks designed to establish a hierarchy of visual components for making comparisons [@cleveland1984graphical]. 
For example, it is more effective to display information on an $x$ or $y$ axis rather than using color in order to reduce the visual effort necessary to make numerical comparisons.
@cleveland1985graphical found that assessing the position of points along an axis is easier than determining the slope of a line. \svp{XXX we probably could include a few more papers and discussions here - the Cleveland 1984 and 1985 papers are based on the same study and are basically duplicates.}
The results of these cognitive tasks provide some consistent guidance for chart design; however, other methods of visual testing can further evaluate design choices and help us understand cognitive biases related to the evaluation of statistical charts.

>>>>

Much of the research on the perception of visual features in charts has been conducted in psychophysics and tests for accuracy and quantitative comparisons when understanding a plot. 
Cleveland and McGill conducted a series of cognitive tasks designed to establish a hierarchy of visual components for making comparisons [@cleveland1984graphical]. 
For example, it is more effective to display information on an $x$ or $y$ axis rather than using color in order to reduce the visual effort necessary to make numerical comparisons.
@cleveland1985graphical found that assessing the position of points along an axis is easier than determining the slope of a line. \ear{The design choices we make can affect a viewers ability to judge relationships between $x$ and $y$ coordinates in a scatter plot. For instance, when the data appear dense, viewers tend to overestimate the magnitude of the correlation coefficient} [@cleveland1982variables; @lauer1989density]. @cleveland1993visualizing. \ear{provides an argument for displaying cyclical patterns with an aspect ratio which sets the curve closet to 45$^{\circ}$.}
<!-- \svp{XXX we probably could include a few more papers and discussions here - the Cleveland 1984 and 1985 papers are based on the same study and are basically duplicates.} -->

+ \ear{Gestalt principles of perceptual organization play a crucial role in data extraction [@kosslyn2006graph]: for example, data points that are closer in space are more likely to be perceived as grouped [@ciccione2020grouping], and vertical and horizontal lines are easier to discriminate than oblique ones [@appelle1972perception].}

The results of these cognitive tasks provide some consistent guidance for chart design; however, other methods of visual testing can further evaluate design choices and help us understand cognitive biases related to the evaluation of statistical charts.

### 2022-02-04 Eye-Fitting-Straight-Lines-in-the-Modern-Era.tex
just a few more edits to go.

Much of the research on the perception of visual features in charts has
been conducted in psychophysics and tests for accuracy and quantitative
comparisons when understanding a plot. Cleveland and McGill conducted a
series of cognitive tasks designed to establish a hierarchy of visual
components for making comparisons \citep{cleveland1984graphical}. For
example, it is more effective to display information on an \(x\) or
\(y\) axis rather than using color in order to reduce the visual effort
necessary to make numerical comparisons. \citet{cleveland1985graphical}
found that assessing the position of points along an axis is easier than
determining the slope of a line.
{\textcolor{RedOrange}{XXX we probably could include a few more papers and discussions here - the Cleveland 1984 and 1985 papers are based on the same study and are basically duplicates.}}

\begin{itemize}
\item
  {\textcolor{blue}{Viewers often mentally exaggerate the magnitude of correlations in scatterplots when the data appear dense [@cleveland1982variable; @lauer1989density].}}
\item
  {\textcolor{blue}{Similarly, the aspect ratio of a graph can have an influence on the patterns that viewers identify. @cleveland1993visualizing, for example, argues that viewers can most easily detect cyclical patterns when an aspect ratio that makes the curve closest to 45◦ is used.}}
\item
  {\textcolor{blue}{Gestalt principles of perceptual organization play a crucial role in data extraction [@kosslyn2006graph]: for example, data points that are closer in space are more likely to be perceived as grouped [@ciccione2020grouping], and vertical and horizontal lines are easier to discriminate than oblique ones [@appelle1972perception].}}
\end{itemize}

>>>>

Much of the research on the perception of visual features in charts has
been conducted in psychophysics and tests for accuracy and quantitative
comparisons when understanding a plot. Cleveland and McGill conducted a
series of cognitive tasks designed to establish a hierarchy of visual
components for making comparisons \citep{cleveland1984graphical}. For
example, it is more effective to display information on an \(x\) or
\(y\) axis rather than using color in order to reduce the visual effort
necessary to make numerical comparisons. \citet{cleveland1985graphical}
found that assessing the position of points along an axis is easier than
determining the slope of a line.
{\textcolor{blue}{The design choices we make can affect a viewers ability to judge relationships between $x$ and $y$ coordinates in a scatter plot. For instance, when the data appear dense, viewers tend to overestimate the magnitude of the correlation coefficient}}
\citep{cleveland1982variables, lauer1989density}.
\citet{cleveland1993visualizing}.
{\textcolor{blue}{provides an argument for displaying cyclical patterns with an aspect ratio which sets the curve closet to 45$^{\circ}$.}}

\begin{itemize}
\item
  {\textcolor{blue}{Gestalt principles of perceptual organization play a crucial role in data extraction [@kosslyn2006graph]: for example, data points that are closer in space are more likely to be perceived as grouped [@ciccione2020grouping], and vertical and horizontal lines are easier to discriminate than oblique ones [@appelle1972perception].}}
\end{itemize}
